---
title: "Connor Practice 2"
author: "Connor McCaughan"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include = FALSE}
library(tidycensus)
library(tidyverse)
library(dplyr)
library(scales)
```

```{r get_big_cities, include = FALSE}
years <- c(2005:2019, 2021:2023)
get_big_city <- function(year) {
  pop <- get_acs(
    geography = 'place',
    variables = c(pop = 'B01003_001'),
    year = year,
    survey = 'acs1'
  )
  pop %>%
    filter(estimate > 700000) %>%
             select(GEOID, NAME, population = estimate) %>%
             mutate(year = year)
}
big_cities <- map_df(years, get_big_city)
```

```{r get_tenure, include = FALSE} 
owner <- 'B25003_002'
renter <- 'B25003_003'
total <- 'B25003_001'
get_tenure <- function(year) {
  tenure <- get_acs(
    geography = 'place',
    variables = c(renter_occupied = renter, owner_occupied = owner, total_occupied = total),
    year = year,
    survey = 'acs1'
  ) %>%
    select(GEOID, variable, estimate) %>%
    pivot_wider(names_from = variable, values_from = estimate) %>%
    mutate(
      year = year,
      owner_pct = owner_occupied/total_occupied,
      renter_pct = renter_occupied/total_occupied
    )
}

tenure_all <- map_df(years, get_tenure)
tenure_big <- tenure_all %>%
  inner_join(big_cities %>% select(GEOID, year), by = c("GEOID", 'year'))
```

```{r get_median_gross_rent, include = FALSE}
med_gross_rent <- 'B25064_001'
```

```{r get_med_monthly_homeowner_costs, include = FALSE}
monthly_ho_cost <- 'B25088_002'
```

```{r get_costs, include = FALSE} 
get_costs <- function(year) {
  costs <- get_acs(
    geography = 'place',
    variables = c(rent = med_gross_rent,
                  owner_cost = monthly_ho_cost),
    year = year,
    survey = 'acs1'
  ) %>%
    select(GEOID, variable, estimate) %>%
    pivot_wider(names_from = variable, values_from = estimate) %>%
    mutate(year = year)
}

costs_all <- map_df(years, get_costs)
costs_big <- costs_all %>%
  inner_join(big_cities %>%
               select(GEOID, year),
             by = c("GEOID", "year"))
```

```{r aggregate_years, include = FALSE} 
tenure_summary <- tenure_big %>%
  group_by(year) %>%
  summarize(
    owner_pct = median(owner_pct, na.rm = TRUE),
    renter_pct = median(renter_pct, na.rm = TRUE)
  )

cost_summary <- costs_big %>%
  group_by(year) %>%
  summarize(
    median_rent = median(rent, na.rm = TRUE),
    median_owner_cost = median(owner_cost, na.rm = TRUE)
  )

final_data <- tenure_summary %>%
  left_join(cost_summary, by = 'year')

```

```{r scaling} 
scale_factor <- max(final_data$median_rent, final_data$median_owner_cost)/
                max(final_data$owner_pct + final_data$renter_pct)
  

final_data <- final_data %>%
  mutate(
    rent_scaled = median_rent/scale_factor,
    owner_cost_scaled = median_owner_cost/scale_factor
    )

final_long <- final_data %>%
  pivot_longer(
    cols = c(owner_pct, renter_pct),
    names_to = "tenure_type",
    values_to = 'percent'
  ) %>%
  mutate(tenure_type = recode(tenure_type,
                              owner_pct = "Owner",
                              renter_pct = "Renter"))

final_long <- final_long %>%
  mutate(percent = percent * 100)

```

```{r mortgage} 
ggplot(final_long, aes(x = factor(year), y = percent, fill = tenure_type)) +
  geom_col() +
  
  geom_line(
    data = final_data %>% mutate(renter_pct = renter_pct * 100),
    aes(x = factor(year), y = renter_pct, color = "% Renter"),
    size = 1,
    inherit.aes = FALSE,
    fill = NA
  ) +
  
  scale_y_continuous(name = "Occupancy Type (%)") +
  theme_minimal() +
  scale_fill_manual(values = c("Owner" = "green3", "Renter" = "darkgreen")) +
  scale_color_manual(values = c("% Renter" = "red")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

 
```
##Mortage vs Rental Trends
This graph shows the proportion of owner-occupied hosuing units to renter-occupied housing units with overlays of median rent price and median monthly homeowner costs from 2005 to 2023.

```{r line_graphs}
final_long <- final_data %>%
  select(year, median_rent, median_owner_cost) %>%
  pivot_longer(
    cols = -year,
    names_to = "variable",
    values_to = "value"
  )
variable_labels <- c(
  median_rent = "Median Rent ($)",
  median_owner_cost = "Median Owner Cost ($)"
)
variable_colors <- c(
  median_rent = 'red',
  median_owner_cost = 'orange'
)
ggplot(final_long, aes(x=year, y=value, color = variable)) +
  # geom_line(size = 1) +
  # geom_point(size = 2) +
  geom_smooth() +
  scale_color_manual(values = variable_colors, labels = variable_labels) +
  scale_y_continuous(
    name = "Dollars ($)",
    labels = scales::dollar
  ) +
  labs(
    x = "Year",
    title = "Housing Tenure and Costs in Big Cities (Pop > 700,000)"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.title = element_blank()
  )
```

######################### Real Practice ###########################################################

#Rent vs. Owner costs nationally with filters for cities and years?
  median monthly cost of rent vs median monthly cost of owning a home (mortgage + avg repairs, etc)
    compare national median vs big cities

#Rent vs. Homeownership Rate Trend with filter for cities
  As rents rise, do cities become less owner-dominated?
    plot Homeownership rate, renter share, and median rent

#inflation adjusted rent trends?
  add inflation data to the above visualizations to get a better view of what the trends really      look like

#add vertical lines at 2008, 2014, and 2020
















































































