---
title: "ShinyPractice_CostlyLiving"
author: "Connor McCaughan"
date: "`r Sys.Date()`"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(tidyverse, tidycensus, scales, shiny, dplyr, ggplot2, lubridate, plotly)
```

```{r census_key, include = FALSE}
census_api_key("3256bc5e0343cacd70f9f681995f33050910c491", install = TRUE, overwrite = TRUE)
readRenviron('~/.Renviron')
```

#Rent vs. Owner costs nationally with filters for cities and years?
  median monthly cost of rent vs median monthly cost of owning a home (mortgage + avg repairs, etc)
    compare national median vs big cities

#Rent vs. Homeownership Rate Trend with filter for cities
  As rents rise, do cities become less owner-dominated?
    plot Homeownership rate, renter share, and median rent

#inflation adjusted rent trends?
  add inflation data to the above visualizations to get a better view of what the trends really      look like

#add vertical lines at 2008, 2014, and 2020


```{r get_big_cities, include = FALSE}
years <- c(2005:2019, 2021:2023)
get_big_city <- function(year) {
  pop <- get_acs(
    geography = 'place',
    variables = c(pop = 'B01003_001'),
    year = year,
    survey = 'acs1'
  )
  pop %>%
    filter(estimate > 700000) %>%
             select(GEOID, NAME, population = estimate) %>%
             mutate(year = year)
}
big_cities <- map_df(years, get_big_city)
```


```{r get_costs, include = FALSE} 
med_gross_rent <- 'B25064_001'
monthly_ho_cost <- 'B25088_002'

get_costs <- function(year) {
  costs <- get_acs(
    geography = 'place',
    variables = c(rent = med_gross_rent,
                  owner_cost = monthly_ho_cost),
    year = year,
    survey = 'acs1'
  ) %>%
    select(GEOID, variable, estimate) %>%
    pivot_wider(names_from = variable, values_from = estimate) %>%
    mutate(year = year)
}

costs_all <- map_df(years, get_costs)
# 
# costs_big <- costs_all %>%
#   inner_join(big_cities %>%
#                select(GEOID, year),
#              by = c("GEOID", "year"))

national_costs <- costs_all %>%
  group_by(year) %>%
  summarize(national_median_rent = median(med_gross_rent, na.rm = TRUE),
            national_median_owner_cost = median(monthly_ho_cost, na.rm = TRUE))
```

```{r data_cache, include = FALSE}
write_rds(costs_all, "/Users/connor/Costly_Living/ConnorPractice_files/costs_all.rds")
write_rds(national_costs, "/Users/connor/Costly_Living/ConnorPractice_files/national_costs.rds")
```

```{r plot_costs}

```































